---
title: "Practical Macheine Learning Project"
author: "me"
date: "14/10/2021"
output:
  pdf_document: default
  html_document: default
---

**Overview**
*When loading the data it was evident that there was a large amount of data that would not be useful for the analysis. Therefore, I started by removing the values of NA, values close to 0 and the columns which were not going to contribute. Then I separated the training data into train_1 and train_2 so I could create a model with train_1 and test the accuracy with train_2. To determine the most accurate predictive model I tested the accuracy of RF, GBM, and LDA and will use the most accurate for the testing data.*

```{r data}
testing <- read.csv("~/Desktop/pml-testing.csv")
training <- read.csv("~/Desktop/pml-training.csv")
training <- training[ , colSums(is.na(training))==0]
training <- training[,-c(1:7)]
testing <- testing[ , colSums(is.na(testing))==0]

```



```{r split data into train and test sets}
library(caret)
set.seed(1234)
inTrain <- createDataPartition(y=training$classe, p=0.7, list = FALSE)
train_1 <- training[inTrain, ]
train_2 <- training[-inTrain, ]
train_1 <- train_1[,-(nearZeroVar(train_1))]
train_2 <- train_2[,-(nearZeroVar(train_2))]
dim(train_1)
dim(train_2)
```

```{r random forest}
library(AppliedPredictiveModeling)
library(gbm)
library(lda)
library(randomForest)
set.seed(1234)
control <- trainControl(method="cv", number=3, verboseIter=FALSE)
rf_model <- train(classe ~ ., data=train_1, method="rf", trControl=control)
rf_model$finalModel
```

```{r random forest predict}
rf_pred <- predict(rf_model, newdata=train_2)
rf_mat <- confusionMatrix(rf_pred, factor(train_2$classe))$overall['Accuracy']
rf_mat
```
*The accuracy for the random forest model is 0.9934 with an out-of-sample error of 0.0066*


```{r gbm}
gbm_model <- train(classe ~ ., data=train_1, method="gbm", trControl=control)
gbm_model$finalModel
gbm_pred <- predict(gbm_model, newdata=train_2)
gbm_mat <- confusionMatrix(gbm_pred, factor(train_2$classe))$overall['Accuracy']
gbm_mat
```
*The accuracy for the gbm model is 0.9670 with an out-of-sample error of 0.0330.*

```{r lda}
lda_model <- train(classe ~ ., data=train_1, method="lda", trControl=control)
lda_model$finalModel
lda_pred <- predict(lda_model, newdata=train_2)
lda_mat <- confusionMatrix(lda_pred, factor(train_2$classe))$overall['Accuracy']
lda_mat
```
*The accuracy for the lda model is 0.6960 with an out-of-sample error of 0.3040.*

**Of the three different prediction models: RF, GBM, and LDA the most accurate prediction model is Random Forest with an accuracy of 0.993373. This is the model which will be used for the Course Project Prediction Quiz.**

